---
title: "Associations Rules"
output: html_notebook
---
Source : http://www.rdatamining.com/examples/association-rules

Read Titanic Data:
```{r}
load("/Users/David/Downloads/titanic.raw.rdata")
str(titanic.raw)
```
Association Rule Mining
```{r}
library(arules)
# find association rules with default settings
rules <- apriori(titanic.raw)
inspect(rules)
```
We then set rhs=c("Survived=No", "Survived=Yes") in appearance to make sure that only "Survived=No" and "Survived=Yes" will appear in the rhs of rules.
```{r}
# rules with rhs containing "Survived" only
rules <- apriori(titanic.raw, parameter = list(minlen=2, supp=0.005, conf=0.8), appearance = list(rhs=c("Survived=No", "Survived=Yes"), default="lhs"), control = list(verbose=F))
rules.sorted <- sort(rules, by="lift")
inspect(rules.sorted)
```
Pruning Redundant Rules

In the above result, rule 2 provides no extra knowledge in addition to rule 1, since rules 1 tells us that all 2nd-class children survived. Generally speaking, when a rule (such as rule 2) is a super rule of another rule (such as rule 1) and the former has the same or a lower lift, the former rule (rule 2) is considered to be redundant. Below we prune redundant rules.
```{r}
# find redundant rules
subset.matrix <- is.subset(rules.sorted, rules.sorted)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1
which(redundant) 
# [1] 2 4 7 8
# remove redundant rules
rules.pruned <- rules.sorted[!redundant]
inspect(rules.pruned)
```
Visualizing Association Rules

Package arulesViz supports visualization of association rules with scatter plot, balloon plot, graph, parallel coordinates plot, etc.
```{r}
library(arulesViz)
plot(rules)
plot(rules, method="graph", control=list(type="items"))
plot(rules, method="paracoord", control=list(reorder=TRUE))
```