---
title: "kNN with NCCTG Lung Cancer Data"
output: html_notebook
---
Source : https://vincentarelbundock.github.io/Rdatasets/doc/survival/cancer.html
Description:

inst:	Institution code
time:	Survival time in days
status:	censoring status 1=censored, 2=dead
age:	Age in years
sex:	Male=1 Female=2
ph.ecog:	ECOG performance score (0=good 5=dead)
ph.karno:	Karnofsky performance score (bad=0-good=100) rated by physician
pat.karno:	Karnofsky performance score as rated by patient
meal.cal:	Calories consumed at meals
wt.loss:	Weight loss in last six months

Load data:
```{r}
library(survival)
data(cancer)
summary(cancer)
str(cancer)
```
Show Data:
```{r}
cancer
```
Remove NA Values
```{r}
clearedData <- cancer[complete.cases(cancer),]
# add target column
table(clearedData$status_result)
clearedData$status_result <- factor(clearedData$status, levels = c(1, 2), labels = c("censored", "dead"))
```
Normalize Data for PCA and kNN:
```{r}
# normalize
normalizedData <- scale(clearedData[,1:10])
summary(normalizedData)
```
PCA:
```{r}
pca <- prcomp(normalizedData, center=TRUE, scale. = TRUE)
biplot(pca, cex=0.6, expand=1)
```
Split in Training and Testdata:
```{r}
train_data <- normalizedData[1:55,]
test_data <- normalizedData[55:167,]
# labels
lbl_train_data <- clearedData[1:55, 11]
lbl_test_data <- clearedData[55:167, 11]
# installs
# install.packages("class")
# library(class)
predict <- knn(train = train_data, test=test_data, cl=lbl_train_data, k=10 )
# install.packages("gmodels")
library(gmodels)
CrossTable(x = lbl_test_data, y = predict, prop.chisq = FALSE)
```