---
title: "RegressionVSClasssification"
output: html_notebook
---

We use iris and prestige data:

Load Prestige Data

```{r}
library(car)
```
```{r}
Prestige
```
```{r}
summary(Prestige)
```
```{r}
pairs(main="Prestige Data", Prestige)
```
```{r}
plot.ts(main="Prestige Data", Prestige)
```
```{r}
d <- data.frame(education = Prestige$education, income = Prestige$income, women = Prestige$women, prestige = Prestige$prestige, census = Prestige$census)
m <- cor(d)
library('corrplot')
corrplot(m, method="circle")
```
Education and income has a high influence to the prestige !!!
Prepase Data Prestige:
```{r}
testidx_pre <- which(1:nrow(Prestige)%%4==0)
```
```{r}
prestige_train <- Prestige[-testidx_pre,]
```
```{r}
prestige_test <- Prestige[testidx_pre,]
```
Prepare Data IRIS:
```{r}
pairs(main="Iris Data", iris)
```
```{r}
plot.ts(main="Iris Data", iris)
```
Create build index:
```{r}
testidx <- which(1:length(iris[,1])%%5==0)
```
separate training data from test:
```{r}
iristrain <- iris[-testidx,]
```
```{r}
iristest <- iris[testidx,]
```
--REGRESSION--
It based on assumption that a linear relationship exists between the input and output variables.
```{r}
model <- lm(prestige~., data=prestige_train)
```
Use model for repdiction the test data
```{r}
prediciton <- predict(model, newdata=prestige_test)
```
Check coorelation with current result
```{r}
cor(prediciton, prestige_test$prestige)
summary(model)
```
```{r}
# curve(4.201e+00*x^5 + 1.150e-03*x^4 + 3.630e-02*x^3 + 1.865e-03*x^2 + 1.131e+01*x^1 + 1.987e+00*x^0 ,from=1, to=200,, xlab="x", ylab="prestige")
curve(4.201e+00*x + 1.150e-03,from=1, to=100, xlab="x", ylab="prestige")
```
```{r}
plot(model)
```
```{r}
prediciton
```

